---
title: "Formatting site coordinates and associated documentation/checks"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---
## Overview
All sites need to be correctly transformed to a shared coordinate system (WGS1984). It's also crucial that sites are correctly indexed to correspond to sites in PECBMS count dataframes (note that for France and UK the datasets were sent in a format with all the data in a shared dataframe already). Information used in decision making is logged, and a few simple checks are carried out. Finally, each country is plotted. 

```{r, message=F, echo=F, warning=F}
library(data.table); library(xlsx); library(dplyr); library(rgdal); library(ggplot2); 
require(maptools); require(mapview); require(gplots)

# helper functions
coordCheck <- function(count, coord) all(unique(count) %in% unique(coord))
f2n <- function(x) as.numeric(as.character(x))
createMaps <- function(x){
        leaflet(data = x) %>% 
            addProviderTiles("OpenStreetMap", group = "OpenStreetMap") %>% 
            addCircleMarkers(group = "sites",
                             lng = coordinates(x)[, 1],
                             lat = coordinates(x)[, 2],
                             radius=5, 
                             color="red",
                             stroke = FALSE, fillOpacity = 0.8,
                             data=x)
}


# Country shape files
# Country polygons obtained here:
# http://ec.europa.eu/eurostat/web/gisco/geodata/reference-data/administrative-units-statistical-units/
countries <- readShapePoly("../files/countryShapeFiles_hiRes/CNTR_2014_03M_SH/Data/worldMap_subset_countriesWithData_clippedToEurope_10kmBuffer.shp")
proj4string(countries) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

# count and coordinate dataframes
fnames_count <- list.files("../recievedFiles/PECBMS_originalFiles/counts/csv/", 
                           full.names=T)

fnames_coord <- list.files("../recievedFiles/PECBMS_originalFiles/coords/", 
                           full.names=T)

# scheme key
schemeKey <- data_frame(country = c("Belgium_Brussels", "Belgium_Wallonia", "Czech",
                                     "Denmark", "Estonia", "Finland", "France", "Germany_East", 
                                     "Germany_West", "Netherlands", "Spain", "Spain_Catalonia", 
                                     "Sweden_New", "Sweden_Old", "UK"), 
                         schemeID = c("BEL.B", "BEL.W", "CZE.0", "DEN.0", "EST.0", 
                                      "FIN.0", "FRA.0", "GER.0", "GER.0", 
                                      "NET.0", "SPA.0", "SPA.C", "SWE.N", "SWE.O", "UKI.0"))
# list to catch each country's coords
cleanedCoords <- list()
```

# proj4 coordinate systems
```{r proj4}
# coordinate systems
LamII <- CRS("+proj=lcc +lat_1=46.8 +lat_0=46.8 +lon_0=0 +k_0=0.99987742 +x_0=600000 +y_0=2200000 +a=6378249.2 +b=6356515 +towgs84=-168,-60,320,0,0,0,0 +pm=paris +units=m +no_defs")

ETRS89 <- CRS("+proj=longlat +ellps=GRS80 +no_defs")

WGS84 <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

LamBelge72 <- CRS("+proj=lcc +lat_1=51.16666723333333 +lat_2=49.8333339 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.013 +y_0=5400088.438 +ellps=intl +towgs84=106.869,-52.2978,103.724,-0.33657,0.456955,-1.84218,1 +units=m +no_defs")

UTM32 <- CRS("+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")

KKJ <- CRS("+proj=longlat +ellps=intl +towgs84=-96.062,-82.428,-121.753,-4.801,-0.345,1.376,1.496 +no_defs")

AMERSFRT <- CRS("+proj=sterea +lat_0=52.15616055555555 +lon_0=5.38763888888889 +k=0.9999079 +x_0=155000 +y_0=463000 +ellps=bessel +units=m +no_defs")

ESTONIA92 <- CRS("+proj=lcc +lat_1=59.33333333333334 +lat_2=58 +lat_0=57.51755393055556 +lon_0=24 +x_0=500000 +y_0=6375000 +ellps=GRS80 +towgs84=0.055,-0.541,-0.185,0.0183,-0.0003,-0.007,-0.014 +units=m +no_defs")
```


# Belgium: Brussels
Straight forward

- single site ID column in coordinate df
- all 'Site' values in count df are found in corresponding column of coord df
- format from Lambert Belge II to WGS84
```{r, message=F, echo=F}
i <- 1
# straight forward- just a single site ID column
print(fnames_count[[i]])
print(fnames_coord[[i]])

counts_i <- fread(fnames_count[[i]], showProgress = F)
coords_i <- read.xlsx2(fnames_coord[[i]], 1)

paste0("all count sites in coord sites? ", all(counts_i$Site %in% coords_i$site))
paste0("any duplicated sites? ", any(duplicated(coords_i$site)))


# format to WGS84
clean_i <- coords_i %>% 
    mutate(country = "Belgium_Brussels") %>%
    left_join(., schemeKey) %>%
    mutate(siteID = paste0(schemeID, ":", site))

# missing coord?
missing_coord <- clean_i %>%
    summarise(n_missing = nrow(coords_i) - n(), p_missing = (n_missing/nrow(coords_i)))
print("missing coord?")
print(missing_coord)

# format WGS84
sp_i <- clean_i %>%
    select(x = CENTREX, y = CENTREY) %>%
    mutate(x=f2n(x), y=f2n(y)) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = LamBelge72)

trans_i <- spTransform(sp_i, WGS84)

cleanedCoords[[i]] <- coordinates(trans_i) %>%
    as_data_frame %>%
    select(lon_WGS84=1, lat_WGS84=2) %>%
    bind_cols(clean_i, .) %>%
    select(schemeID, siteID, lon_WGS84, lat_WGS84)

# plot
forPlot_i <- cleanedCoords[[i]] %>%
    select(x = lon_WGS84, y = lat_WGS84) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = WGS84)

createMaps(forPlot_i)
```

# Belgium: Wallonia region
Straight forward

- No ambiguity in site column: single site ID column in coordinate df
- all 'Site' values in count df are found in corresponding column of coord df
- Already in WGS84 format
```{r, message=F, echo=F}
i <- 2
# straight forward- just a single site ID column
print(fnames_count[[i]])
print(fnames_coord[[i]])

counts_i <- fread(fnames_count[[i]], showProgress = F)
coords_i <- read.xlsx2(fnames_coord[[i]], 1)

paste0("all count sites in coord sites? ", coordCheck(counts_i$Site, coords_i$plot_ID))
paste0("any duplicated sites? ", any(duplicated(coords_i$plot_ID)))

# Already formatted to WGS84
clean_i <- coords_i %>% 
    mutate(country = "Belgium_Wallonia") %>%
    left_join(., schemeKey) %>%
    mutate(siteID = paste0(schemeID, ":", plot_ID)) %>%
    mutate(lon_WGS84 = f2n(X1), lat_WGS84 = f2n(Y1)) %>%
    # Wallonia has the additional complication that 243 sites have (0,0) coordinates
    # (out of 4199).
    ## Remove these.
    filter(!(lon_WGS84 == 0 & lat_WGS84==0))

# missing coord?
missing_coord <- clean_i %>%
    summarise(n_missing = nrow(coords_i) - n(), p_missing = (n_missing/nrow(coords_i)))
print("missing coord?")
print(missing_coord)

cleanedCoords[[i]] <- clean_i %>%
    select(schemeID, siteID, lon_WGS84, lat_WGS84)

# plot
forPlot_i <- cleanedCoords[[i]] %>%
    select(x = lon_WGS84, y = lat_WGS84) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = WGS84)

createMaps(forPlot_i)
```

# Czech Republic
Straight forward

- No ambiguity in site column: single site ID column in coordinate df
- all 'Site' values in count df are found in corresponding column of coord df
- note: have elected to use municipality coordinates,on the basis these are quite 
accurate anyway (most accurate to 2km)
- Already in WGS84
```{r, message=F, echo=F}
i <- 3
# straight forward- just a single site ID column
print(fnames_count[[i]])
print(fnames_coord[[i]])

counts_i <- fread(fnames_count[[i]], showProgress = F)
coords_i <- read.xlsx(fnames_coord[[i]], 1)

paste0("all count sites in coord sites? ", coordCheck(counts_i$Site, coords_i$Site))
paste0("any duplicated sites? ", any(duplicated(coords_i$Site)))

clean_i <- coords_i %>% 
    mutate(country = "Czech") %>%
    left_join(., schemeKey) %>%
    mutate(siteID = paste0(schemeID, ":", Site)) %>%
    mutate(lon_WGS84 = f2n(municipality_LON), lat_WGS84 = f2n(municipality_LAT))

# missing coord?
missing_coord <- clean_i %>%
    summarise(n_missing = nrow(coords_i) - n(), p_missing = (n_missing/nrow(coords_i)))
print("missing coord?")
print(missing_coord)

cleanedCoords[[i]] <- clean_i %>%
    select(schemeID, siteID, lon_WGS84, lat_WGS84)

# plot
forPlot_i <- cleanedCoords[[i]] %>%
    select(x = lon_WGS84, y = lat_WGS84) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = WGS84)

createMaps(forPlot_i)

```

# Denmark
Straight forward

- No ambiguity in site column: single site ID column in coordinate df
- all 'Site' values in count df are found in corresponding column of coord df
- In UTM zone 32 format
```{r, message=F, echo=F}
i <- 4
# straight forward- just a single site ID column
print(fnames_count[[i]])
print(fnames_coord[[i]])

counts_i <- fread(fnames_count[[i]], showProgress = F)
coords_i <- read.xlsx2(fnames_coord[[i]], 1)

paste0("all count sites in coord sites? ", coordCheck(counts_i$Site, coords_i$RouteID))
paste0("any duplicated sites? ", any(duplicated(coords_i$RouteID)))

# format to WGS84
clean_i <- coords_i %>% 
    mutate(country = "Denmark") %>%
    left_join(., schemeKey) %>%
    mutate(siteID = paste0(schemeID, ":", RouteID), 
           x = Easting, y = Northing,
           x=f2n(x), y=f2n(y)) %>%
    filter(!is.na(x)|!is.na(y)) %>%
    # single site that doesn't have coordinates, but has (0,0) as (N, E). 
    # Other missing coordinate sites actually have blanks for (N, E)
    filter(!RouteID == 3458)

# missing coord?
missing_coord <- clean_i %>%
    summarise(n_missing = nrow(coords_i) - n(), p_missing = (n_missing/nrow(coords_i)))
print("missing coord?")
print(missing_coord)

# format WGS84
sp_i <- clean_i %>%
    select(x, y) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = UTM32)

trans_i <- spTransform(sp_i, WGS84)

cleanedCoords[[i]] <- coordinates(trans_i) %>%
    as_data_frame %>%
    select(lon_WGS84=1, lat_WGS84=2) %>%
    bind_cols(clean_i, .) %>%
    select(schemeID, siteID, lon_WGS84, lat_WGS84)

# plot
forPlot_i <- cleanedCoords[[i]] %>%
    select(x = lon_WGS84, y = lat_WGS84) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = WGS84)

createMaps(forPlot_i)
```

# Estonia
Straight forward

- No ambiguity in site column: single site ID column in coordinate df
- all 'Site' values in count df are found in corresponding column of coord df
- In Estonia'92 coordinate system (see CRS for details). 
- No missing coordinates
```{r, message=F, echo=F}
i <- 5
# straight forward- just a single site ID column
print(fnames_count[[i]])
print(fnames_coord[[i]])

counts_i <- fread(fnames_count[[i]], showProgress = F)
coords_i <- read.xlsx2(fnames_coord[[i]], 1)

paste0("all count sites in coord sites? ", coordCheck(counts_i$Site, coords_i$site))
paste0("any duplicated sites? ", any(duplicated(coords_i$site)))

# format to WGS84
clean_i <- coords_i %>% 
    mutate(country = "Estonia") %>%
    left_join(., schemeKey) %>%
    mutate(siteID = paste0(schemeID, ":", site), 
           x = X_COORD, y = Y_COORD,
           x=f2n(x), y=f2n(y)) %>%
    filter(!is.na(x)|!is.na(y))

# missing coord?
missing_coord <- clean_i %>%
    summarise(n_missing = nrow(coords_i) - n(), p_missing = (n_missing/nrow(coords_i)))
print("missing coord?")
print(missing_coord)

sp_i <- clean_i %>%
    select(x, y) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = ESTONIA92)

trans_i <- spTransform(sp_i, WGS84)

cleanedCoords[[i]] <- coordinates(trans_i) %>%
    as_data_frame %>%
    select(lon_WGS84=1, lat_WGS84=2) %>%
    bind_cols(clean_i, .) %>%
    select(schemeID, siteID, lon_WGS84, lat_WGS84)

# plot
forPlot_i <- cleanedCoords[[i]] %>%
    select(x = lon_WGS84, y = lat_WGS84) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = WGS84)

createMaps(forPlot_i)
```

# Finland
Straight forward

- No ambiguity in site column: single site ID column in coordinate df
- all 'Site' values in count df are found in corresponding column of coord df
- In KKJ coordinate system
- No missing sites
```{r, message=F, echo=F}
i <- 6
# straight forward- just a single site ID column
print(fnames_count[[i]])
print(fnames_coord[[i]])

counts_i <- fread(fnames_count[[i]], showProgress = F)
coords_i <- read.xlsx2(fnames_coord[[i]], 1)

paste0("all count sites in coord sites? ", coordCheck(counts_i$Site, coords_i$Site))
paste0("any duplicated sites? ", any(duplicated(coords_i$site)))

# format to WGS84
clean_i <- coords_i %>% 
    mutate(country = "Finland") %>%
    left_join(., schemeKey) %>%
    mutate(siteID = paste0(schemeID, ":", Site), 
           x = lon, y = kkj.decimal.lat,
           x=f2n(x), y=f2n(y)) %>%
    filter(!is.na(x)|!is.na(y))

# missing coord?
missing_coord <- clean_i %>%
    summarise(n_missing = nrow(coords_i) - n(), p_missing = (n_missing/nrow(coords_i)))
print("missing coord?")
print(missing_coord)

sp_i <- clean_i %>%
    select(x, y) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = KKJ)

trans_i <- spTransform(sp_i, WGS84)

cleanedCoords[[i]] <- coordinates(trans_i) %>%
    as_data_frame %>%
    select(lon_WGS84=1, lat_WGS84=2) %>%
    bind_cols(clean_i, .) %>%
    select(schemeID, siteID, lon_WGS84, lat_WGS84)

# plot
forPlot_i <- cleanedCoords[[i]] %>%
    select(x = lon_WGS84, y = lat_WGS84) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = WGS84)

createMaps(forPlot_i)
```

# Germany: East

- Some ambiguity in site column- several code columns- but the Sites in the count dataframe are all integer valued, wheras routecodes are not. The 'Site in F1 data' column additionally corresponds to the sites in the count df
- all 'Site' values in count df are found in corresponding column of coord df
- In WGS84 already
- No missing coordinates
```{r, message=F, echo=F}
i <- 7
# straight forward- just a single site ID column
print(fnames_count[[i]])
print(fnames_coord[[i]])

counts_i <- fread(fnames_count[[i]], showProgress = F)
coords_i <- read.xlsx2(fnames_coord[[i]], 2)

paste0("all count sites in coord sites? ", coordCheck(counts_i$Site, coords_i$site.in.F1.data))
paste0("any duplicated sites? ", any(duplicated(coords_i$site.in.F1.data)))

# format to WGS84
clean_i <- coords_i %>% 
    mutate(country = "Germany_East") %>%
    left_join(., schemeKey) %>%
    mutate(siteID = paste0(schemeID, ":", site.in.F1.data), 
           x = long.in..shp, y = lat.in..shp,
           x=f2n(x), y=f2n(y)) %>%
    filter(!is.na(x)|!is.na(y))

# missing coord?
missing_coord <- clean_i %>%
    summarise(n_missing = nrow(coords_i) - n(), p_missing = (n_missing/nrow(coords_i)))
print("missing coord?")
print(missing_coord)

cleanedCoords[[i]] <- clean_i %>%
    mutate(lon_WGS84 = x, lat_WGS84 = y) %>%
    select(schemeID, siteID, lon_WGS84, lat_WGS84)

# plot
forPlot_i <- cleanedCoords[[i]] %>%
    select(x = lon_WGS84, y = lat_WGS84) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = WGS84)

createMaps(forPlot_i)
```

# Germany: West

- Some ambiguity in site column- several code cols- but the Sites in the count dataframe are all integer valued, wheras routecodes are not. The 'Site in F1' column additionally corresponds to the sites in the count df
- Additional complication is that I am merging Germany E and Germany W into the same df (there are no methodological differences between schemes, split is just for historical reasons). Therefore, siteIDs have to be unique across both schemes, rather than just internally. Have checked and they are unique. 
- Two sites in the count dataframe are not in the site dataframe. Will just remove these
- An additional 1 site has no coordinates
- Already in WGS84
```{r, message=F, echo=F}
i <- 8
# straight forward- just a single site ID column
print(fnames_count[[i]])
print(fnames_coord[[i]])

counts_i <- fread(fnames_count[[i]], showProgress = F)
coords_i <- read.xlsx2(fnames_coord[[i]], 2)

# check that there is no overlap in site IDs between dataframes: there isn't
coords_GE <- read.xlsx2(fnames_coord[[7]], 2)
paste0("any non-unique ids across German schemes?: ",
       any(coords_GE$site.in.F1.data %in% coords_i$Site.in.F1.data) |
           any(coords_i$Site.in.F1.data %in% coords_GE$site.in.F1.data))

paste0("all count sites in coord sites? ", coordCheck(counts_i$Site, coords_i$Site.in.F1.data))
# which(!unique(counts_i$Site) %in% coords_i$Site.in.F1.data)
paste0("any duplicated sites? ", any(duplicated(coords_i$Site.in.F1.data)))

# format to WGS84
clean_i <- coords_i %>% 
    mutate(country = "Germany_West") %>%
    left_join(., schemeKey) %>%
    mutate(siteID = paste0(schemeID, ":", Site.in.F1.data), 
           x = long.in..shp, y = lat.in..shp,
           x=f2n(x), y=f2n(y)) %>%
    filter(!is.na(x)|!is.na(y))

# missing coord?
missing_coord <- clean_i %>%
    summarise(n_missing = nrow(coords_i) - n(), p_missing = (n_missing/nrow(coords_i)))
print("missing coord?")
print(missing_coord)

cleanedCoords[[i]] <- clean_i %>%
    mutate(lon_WGS84 = x, lat_WGS84 = y) %>%
    select(schemeID, siteID, lon_WGS84, lat_WGS84)

# plot
forPlot_i <- cleanedCoords[[i]] %>%
    select(x = lon_WGS84, y = lat_WGS84) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = WGS84)

createMaps(forPlot_i)
```

# Netherlands

- No ambiguity in site column 
- all 'Site' values in count df are found in corresponding column of coord df
- In Amersfoort coordinate system
- No missing coordinates
```{r, message=F, echo=F}
i <- 9
# straight forward- just a single site ID column
print(fnames_count[[i]])
print(fnames_coord[[i]])

counts_i <- fread(fnames_count[[i]], showProgress = F)
coords_i <- read.xlsx2(fnames_coord[[i]], 1)

paste0("all count sites in coord sites? ", coordCheck(counts_i$Site, coords_i$Site))
paste0("any duplicated sites? ", any(duplicated(coords_i$Site)))

# format to WGS84
clean_i <- coords_i %>% 
    mutate(country = "Netherlands") %>%
    left_join(., schemeKey) %>%
    mutate(siteID = paste0(schemeID, ":", Site), 
           x = X, y = Y,
           x=f2n(x), y=f2n(y)) %>%
    filter(!is.na(x)|!is.na(y))

# missing coord?
missing_coord <- clean_i %>%
    summarise(n_missing = nrow(coords_i) - n(), p_missing = (n_missing/nrow(coords_i)))
print("missing coord?")
print(missing_coord)

sp_i <- clean_i %>%
    select(x, y) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = AMERSFRT)

trans_i <- spTransform(sp_i, WGS84)

cleanedCoords[[i]] <- coordinates(trans_i) %>%
    as_data_frame %>%
    select(lon_WGS84=1, lat_WGS84=2) %>%
    bind_cols(clean_i, .) %>%
    select(schemeID, siteID, lon_WGS84, lat_WGS84)

# plot
forPlot_i <- cleanedCoords[[i]] %>%
    select(x = lon_WGS84, y = lat_WGS84) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = WGS84)

createMaps(forPlot_i)
```

# Spain 

- Have not been able to supply site IDs, but rather have a KML file to 10km grid level. I've converted these from KML to shape in ArcMap, and then formatted in `create_formatSpanishCoordinates.R`. The processing is thus slightly different here than for other countries
- Note that, while 11 sites are dropped, these are all island sites that have a 5x5km coordinate, but not a 10x10km. (can easily see this by querying ArcMap attribute table). Given that I won't include these in analysis anyway, this isn't a problem
```{r, message=F, echo=F}
i <- 10
# straight forward- just a single site ID column
print(fnames_count[[i]])
print(fnames_coord[[i]])

counts_i <- fread(fnames_count[[i]])
coords_i <- readRDS("../recievedFiles/Spain_coordinates/coordinates_WGS84.rds")


paste0("all count sites in coord sites? ", coordCheck(counts_i$Site, coords_i$siteID))
# which(!unique(counts_i$Site) %in% coords_i$Site.in.F1.data)
paste0("any duplicated sites? ", any(duplicated(coords_i$siteID)))

# format to WGS84
clean_i <- coords_i %>% 
    mutate(country = "Spain", schemeID="SPA.0", siteID=paste0(schemeID, ":", siteID)) %>%
    filter(!is.na(lon_WGS84)|!is.na(lat_WGS84))
# missing coord?
missing_coord <- clean_i %>%
    summarise(n_missing = nrow(coords_i) - n(), p_missing = (n_missing/nrow(coords_i)))
print("missing coord?")
print(missing_coord)

cleanedCoords[[i]] <- clean_i %>%
    select(schemeID, siteID, lon_WGS84, lat_WGS84)

# plot
forPlot_i <- cleanedCoords[[i]] %>%
    select(x = lon_WGS84, y = lat_WGS84) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = WGS84)

createMaps(forPlot_i)
```

# Catalonia

- No ambiguity in site column 
- all 'Site' values in count df are found in corresponding column of coord df
- In WGS84 coordinate system
```{r, message=F, echo=F}
i <- 11
# straight forward- just a single site ID column
print(fnames_count[[i]])
print(fnames_coord[[i]])

counts_i <- fread(fnames_count[[i]], showProgress = F)
coords_i <- read.xlsx2(fnames_coord[[i]], 1)

paste0("all count sites in coord sites? ", coordCheck(counts_i$Site, coords_i$SITE))
paste0("any duplicated sites? ", any(duplicated(coords_i$SITE)))

# format to WGS84
clean_i <- coords_i %>% 
    mutate(country = "Spain_Catalonia") %>%
    left_join(., schemeKey) %>%
    mutate(siteID = paste0(schemeID, ":", SITE), 
           x = LONGITUDE, y = LATITUDE,
           x=f2n(x), y=f2n(y)) %>%
    filter(!is.na(x)|!is.na(y))

# missing coord?
missing_coord <- clean_i %>%
    summarise(n_missing = nrow(coords_i) - n(), p_missing = (n_missing/nrow(coords_i)))
print("missing coord?")
print(missing_coord)

cleanedCoords[[i]] <- clean_i %>%
    mutate(lon_WGS84 = x, lat_WGS84 = y) %>%
    select(schemeID, siteID, lon_WGS84, lat_WGS84)

# plot
forPlot_i <- cleanedCoords[[i]] %>%
    select(x = lon_WGS84, y = lat_WGS84) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = WGS84)

createMaps(forPlot_i)
```

# Sweden: New

- No ambiguity in site column 
- all 'Site' values in count df are found in corresponding column of coord df
- "Each route is a 8 km line transect, shaped as a square (2 km x 2 km). The coordinate is the centerpoint of this "square", thus a point where no counting was carried out. But the point nearest to all counts."
- In WGS84 format
```{r, message=F, echo=F}
i <- 12
# straight forward- just a single site ID column
print(fnames_count[[i]])
print(fnames_coord[[i]])

counts_i <- fread(fnames_count[[i]], showProgress = F)
coords_i <- read.xlsx2(fnames_coord[[i]], 1)

paste0("all count sites in coord sites? ", coordCheck(counts_i$Site, coords_i$Route.))
paste0("any duplicated sites? ", any(duplicated(coords_i$Route.)))

# format to WGS84
clean_i <- coords_i %>% 
    mutate(country = "Sweden_New") %>%
    left_join(., schemeKey) %>%
    mutate(siteID = paste0(schemeID, ":", Route.), 
           x = wgs84_lon, y = wgs84_lat,
           x=f2n(x), y=f2n(y)) %>%
    filter(!is.na(x)|!is.na(y))

# missing coord?
missing_coord <- clean_i %>%
    summarise(n_missing = nrow(coords_i) - n(), p_missing = (n_missing/nrow(coords_i)))
print("missing coord?")
print(missing_coord)

cleanedCoords[[i]] <- clean_i %>%
    mutate(lon_WGS84 = x, lat_WGS84 = y) %>%
    select(schemeID, siteID, lon_WGS84, lat_WGS84)

# plot
forPlot_i <- cleanedCoords[[i]] %>%
    select(x = lon_WGS84, y = lat_WGS84) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = WGS84)

createMaps(forPlot_i)
```

# Sweden: Old

- No ambiguity in site column 
- all 'Site' values in count df are found in corresponding column of coord df
- In WGS84 coordinate system
```{r, message=F, echo=F}
i <- 13
# straight forward- just a single site ID column
print(fnames_count[[i]])
print(fnames_coord[[i]])

counts_i <- fread(fnames_count[[i]], showProgress = F)
coords_i <- read.xlsx2(fnames_coord[[i]], 1)

paste0("all count sites in coord sites? ", coordCheck(counts_i$Site, coords_i$ROTUE.))
paste0("any duplicated sites? ", any(duplicated(coords_i$ROTUE.)))

# format to WGS84
clean_i <- coords_i %>% 
    mutate(country = "Sweden_Old") %>%
    left_join(., schemeKey) %>%
    mutate(siteID = paste0(schemeID, ":", ROTUE.), 
           x = wgs84_lon, y = wgs84_lat,
           x=f2n(x), y=f2n(y)) %>%
    filter(!is.na(x)|!is.na(y))

# missing coord?
missing_coord <- clean_i %>%
    summarise(n_missing = nrow(coords_i) - n(), p_missing = (n_missing/nrow(coords_i)))
print("missing coord?")
print(missing_coord)

cleanedCoords[[i]] <- clean_i %>%
    mutate(lon_WGS84 = x, lat_WGS84 = y) %>%
    select(schemeID, siteID, lon_WGS84, lat_WGS84)

# plot
forPlot_i <- cleanedCoords[[i]] %>%
    select(x = lon_WGS84, y = lat_WGS84) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = WGS84)

createMaps(forPlot_i)
```


# France and UK
- Were sent the dataframes directly, which have counts and coordinates in the same dataframe. Therefore need to process slightly differently. 
- UK already in WGS84 format
```{r, message=F, echo=F}
i <- 14
# need to process separately. 
files <- list.files("../recievedFiles/UK/", 
                    pattern = ".csv", full.names=TRUE, recursive=TRUE)
# extract the id file from the list and remove
speciesKey <- fread(files[grepl("btolist", files)], showProgress = F)
files <- files[!grepl("btolist", files)]

getDF <- function(filename) {
    file_i <- fread(filename)
    Euring <- speciesKey[Two.letter.code == unique(file_i$sp),Euring_code]
    file_i[!is.na(count), list(site=square, ETRS89Lat, ETRS89Long)]
}

clean_i <- lapply(files, getDF) %>%
    bind_rows(.) %>%
    unique %>%
    mutate(schemeID = "UKI.0", siteID = paste0(schemeID, ":", site),
           x = ETRS89Long, y = ETRS89Lat,
           x=f2n(x), y=f2n(y))

# missing coord?
print("No missing coord, duplicates etc.")

sp_i <- clean_i %>%
    select(x, y) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = ETRS89)

trans_i <- spTransform(sp_i, WGS84)

cleanedCoords[[i]] <- coordinates(trans_i) %>%
    as_data_frame %>%
    select(lon_WGS84=1, lat_WGS84=2) %>%
    bind_cols(clean_i, .) %>%
    select(schemeID, siteID, lon_WGS84, lat_WGS84)

# plot
forPlot_i <- cleanedCoords[[i]] %>%
    select(x = lon_WGS84, y = lat_WGS84) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = WGS84)

createMaps(forPlot_i)


## France
i <- 15
clean_i <- fread("../recievedFiles/France/stoc_1989_2015.csv", showProgress = F) %>%
    select(SITE, X, Y, XlamII, YlamII) %>%
    unique %>%
    mutate(schemeID = "FRA.0", siteID=paste0(schemeID, ":", SITE), 
           x = X, y = Y,
           x=f2n(x), y=f2n(y))

# missing coord?
print("No missing coord, duplicates etc.")

sp_i <- clean_i %>%
    select(x, y) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = LamII)

trans_i <- spTransform(sp_i, WGS84)

cleanedCoords[[i]] <- coordinates(trans_i) %>%
    as_data_frame %>%
    select(lon_WGS84=1, lat_WGS84=2) %>%
    bind_cols(clean_i, .) %>%
    select(schemeID, siteID, lon_WGS84, lat_WGS84)

# plot
forPlot_i <- cleanedCoords[[i]] %>%
    select(x = lon_WGS84, y = lat_WGS84) %>%
    as.data.frame %>%
    SpatialPoints(., proj4string = WGS84)

createMaps(forPlot_i)
```

# Save
```{r}
coordinates_full <- bind_rows(cleanedCoords)
saveRDS(coordinates_full, "../cleanedFiles/coordinateData_noSpain.rds")
